---
title: "mixcuref Manual"
author: "Changchang Xu"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{mixcuref Manual}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

# Overview

`mixcuref` fits Weibull mixture cure models by maximum likelihood or Firth-type penalized likelihood. The package provides direct estimation, profile-likelihood confidence intervals, nested likelihood-ratio tests, and utilities for separation diagnostics.

The main user-facing functions are:

- `mixcure.penal.est()` for model fitting
- `mixcure.penal.profile.CI()` for profile-likelihood confidence intervals
- `mixcure.penal.1d.nested.lrt()` and `mixcure.penal.2d.nested.lrt()` for nested tests
- `separation.detection()` for sparse-data diagnostics
- `fit_mixture_cure()` as a thin helper used by the meta-layer
- `mixcuref_fit()` as the adapter used by `Mixcuref.Meta`

# Package data

The package includes three real breast cancer datasets:

- `ANNbcBMdat1`: high event rate, univariate example
- `ANNbcBMdat2`: lower event rate, univariate example
- `ANNbcBMdat5`: lower event rate, multivariable example

# Fitting a mixture cure model

A mixture cure model is specified using a standard survival formula `Surv(time, status) ~ covariates`. Covariates are used in both the cure and latency components.

```{r}
library(mixcuref)
library(survival)

data(ANNbcBMdat1)

fit_mle <- mixcure.penal.est(
  Surv(Time, CENS == 1) ~ Her2,
  data = ANNbcBMdat1,
  init = c(1, -0.1, -10, 1, 1),
  pl = FALSE
)

fit_ftpl <- mixcure.penal.est(
  Surv(Time, CENS == 1) ~ Her2,
  data = ANNbcBMdat1,
  init = c(1, -0.1, -10, 1, 1),
  pl = TRUE
)

fit_mle$coefficients
fit_ftpl$coefficients
```

# Multivariable example

```{r}
data(ANNbcBMdat5)

fit5 <- mixcure.penal.est(
  Surv(Time, CENS == 1) ~ Her2 + LuminalA + TN + MENS0 + TUMCT,
  data = ANNbcBMdat5,
  init = c(5, rep(0, 5), -10, rep(0, 5), 0.1),
  pl = TRUE
)

fit5$coefficients
coef(fit5)
vcov(fit5)
```

# Profile-likelihood confidence intervals

```{r}
plci1 <- mixcure.penal.profile.CI(
  Surv(Time, CENS == 1) ~ Her2,
  data = ANNbcBMdat1,
  init = c(1, -0.1, -10, 1, 1),
  pl = FALSE,
  apct = 0.05
)

plci1$coefficients
```

# Nested likelihood-ratio tests

```{r}
lrt1 <- mixcure.penal.1d.nested.lrt(
  Surv(Time, CENS == 1) ~ Her2,
  data = ANNbcBMdat1,
  init = c(1, -0.1, -10, 1, 1),
  pl = FALSE,
  loglik = fit_mle$loglikelihood
)

lrt1
```

# Separation diagnostics

```{r}
sep_diag <- separation.detection(fit5, pl = TRUE, nsteps = 20)
sep_diag
```

# Meta-layer adapters

`mixcuref` can still be used directly, but it also provides helper functions that support the `Mixcuref.Meta` package.

```{r}
fit_from_helper <- fit_mixture_cure(
  incidence = ~ Her2,
  latency = ~ Her2,
  time = ANNbcBMdat1$Time,
  status = ANNbcBMdat1$CENS,
  data = ANNbcBMdat1,
  control = list(pl = FALSE)
)

fit_from_helper
```

```{r}
# spec is a model specification object created by Mixcuref.Meta
# mixcuref_fit(spec, control = list(pl = FALSE))
```

# Notes

- `pl = FALSE` requests standard maximum likelihood.
- `pl = TRUE` requests Firth-type penalized likelihood.
- Initial values must be supplied in the order: cure coefficients, survival coefficients, Weibull shape parameter.
- For categorical covariates with more than two categories, create dummy variables before fitting.

# Session information

```{r, eval=T, echo=F}
pkgs <- c(
#  "mixcuref",
  "survival",
  "abind",
  "R.utils"
)

desc_field <- function(pkg, field) {
  x <- tryCatch(utils::packageDescription(pkg, fields = field), error = function(e) NA)
  if (is.null(x) || is.na(x)) "" else x
}

pkg_table <- data.frame(
  Package = pkgs,
  Version = sapply(pkgs, function(x) as.character(packageVersion(x))),
  Built   = sapply(pkgs, desc_field, field = "Built"),
  Remote  = sapply(pkgs, desc_field, field = "RemoteSha"),
  row.names = NULL,
  stringsAsFactors = FALSE
)

knitr::kable(pkg_table, caption = "Package Versions")

cat("\n\n")

#sessionInfo()
```
